---
title: 'Rumsay, Chapter 3: Organizing quantitative data - Charts and graphs. [Ongoing
  - Lattice graphics]'
author: "Maurizio Murino"
date: "3 March 2016"
output: html_document
---
```{r setoption, cache=TRUE, warning=FALSE, message=FALSE, fig.width=12}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE, fig.width=12)
```

```{r, loadlattice}
library(lattice)
library(xtable)
```

### Exercise 1 - 3/3/'16

You lose information from the data when yu create a histogram. What information is lost?

__Solution:__

You do not know the actual value of the data anymore, but only the group where they fall into.

### Exercise 2 - 3/3/'16

Make a histogram from this dataset of test scores: 72, 79, 81, 80, 63, 62, 89, 90, 50, 78, 87, 97, 55, 69, 97, 87, 88, 99, 76, 78, 5, 77, 88, 90, 81. Would a pie chart be appropriated for this data?

```{r 2.1 dataload}
e2 <- c(72, 79, 81, 80, 63, 62, 89, 90, 50, 78, 87, 97, 55, 69, 97, 87, 88, 99, 76, 78, 65, 77, 88, 90, 81)
```
```{r 2.2 histogram, lattice}
histogram(~e2,
          type = "count",
          xlab = "Test scores",
          main = "Test scores",
          breaks = seq(from = 50, to = 100, by = 5))
max(e2)
```

A pie chart is not appropriate. Pie charts are more useful with categorical data.

### Exercise 3 - 4/3/'16

Survey of 45 home-owners. Of them, 2 do not have a TV, 17 have 1, 22 have 2, 3 have 3, 1 has 4. Make a relative frequency histogram of this data and interpret the result.

```{r 3.1 dataload}
e3 <- data.frame(tv = rep(c(0, 1, 2, 3, 4), c(2, 17, 22, 3, 1)))
sum(e3$tv)
```

```{r 3.2 histogram}
histogram(~e3,
          xlab = "TVs",
          main = "Distribution of 73 Tvs between 45 owners",
          breaks = seq(from = 0, to = 4, by = 1))
```

### Exercise 4 - 6/3/'16

Think a loaded die. Make it roll several times. 

```{r 4.1 dataload}
e4 <- data.frame(n = rep(c(1, 2, 3, 4, 5, 6), c(10, 15, 5, 15, 5, 20)))
histogram(~n, data = e4,
          type = "count",
          main = "260 die rolls",
          xlab = "Die result",
          breaks = seq(from = 0, to = 6, by = 1))
```

__A. Make the relative frequency histogram of these results.__

```{r 4.2 plot1}
histogram(~n, data = e4,
          main = "Relative requencies - 260 die rolls",
          xlab = "Die result",
          breaks = seq(from = 0, to = 6, by = 1))
```

__B. You can make a relative frequency histogram from a frequency histogram; can you go to the other direction?__

Well, it is possible, but only by arbitrarily determining a sample size, that is unknow. Hence, it is no really useful.