---
title: 'Rumsay, Chapter 6: Measures of relative standing and the normal distribution.
  [Ongoing]'
author: "Maurizio Murino"
date: "12 March 2016"
output: html_document
---

```{r setoption, cache=TRUE, warning=FALSE, message=FALSE, fig.width=12}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE, fig.width=12)
```

```{r, loadlattice}
library(EasieR)
library(knitr)
```

### Exercise 1 - 12/3/'16

What do you guess are the standard deviations of the two distributions in the figure below?

```{r 1.1 dataload1, echo=FALSE}

set.seed(123)
e1a <- seq(10, 70,length = 10000)
e1b <- dnorm(e1a, mean = 40, sd = 8)
```

```{r plot 1}
plote1 <- xyplot(e1b ~ e1a,
       type = "l",
       horizontal = FALSE,
       scales = list(y = list(at = NULL)),
       main = "Normal distribution, mean: 40, ds: 8, range: 10-70",
       panel = function(x, ...) {
               panel.xyplot(x, ...)
               vlines <- c(32, 40, 48)
               panel.abline( v = vlines, lty = 3)
       })
```     

```{r dataload 2, echo=FALSE}
set.seed(123)
e1c <- seq(10, 70,length = 10000)
e1d <- dnorm(e1c, mean = 40, sd = 4)
```

```{r plot2}
plote2 <- xyplot(e1d ~ e1c,
       type = "l",
       horizontal = FALSE,
       scales = list(y = list(at = NULL)),
       main = "Normal distribution, mean: 40, ds: 4, range: 10-70",
       panel = function(x, ...) {
               panel.xyplot(x, ...)
               vlines <- c(36, 40, 44)
               panel.abline( v = vlines, lty = 3)
       })

## position: (a, b, c, d) b: height from the top; d: height from the bottom

print(plote1, position=c(0, .5, 1, 1), more=TRUE)
print(plote2, position=c(0, 0, 1, .5))
```

__Solution:__ 

In the second plot, the standard deviation is slighly smaller than the first's.

### Exercise 2 - 15/3/'16

Draw a picture of a normal distribution with mean 70 and standard deviation 5.

```{r2.1  data load, echo=FALSE}

e2a <- seq(0, 140, length = 10000)
e2b <- dnorm(e2a, 70, 15)                  
```

```{r2.2 plot}
xyplot(e2b ~ e2a,
       type = "l",
       main = "Normal distribution, mean 70",
       scales = list(y = list(at = NULL), x = list(at = seq(0, 140, 10))),
       panel = function(x, ...){
               panel.xyplot(x, ...)
               vmean <- 70
               panel.abline(v = vmean, lty = 4)
               
       })
```

### Exercise 3 - 24/3/'16

Draw a picture with two normal distributions. Give each a mean of 70, with standard deviations of 5 and 10. How do the distributions differ?

```{R 3.1 plots}
e3a <- seq(40, 100, length = 1000)
e3b <- dnorm(e3a, 70, 5)
plot_e3a <- xyplot(e3b ~ e3a,
                   type = "l",
                   main = "Plot 1 - Mean 70, sd 5")

e3c <- seq(40, 100, length = 1000)
e3d <- dnorm(e3a, 70, 10)
plot_e3b <- xyplot(e3d ~ e3c,
                   type = "l",
                   main = "Plot 2 - Mean 70, sd 10",
                   ylim = range(e3b))
                   

print(plot_e3a, position = c(0, .5, 1, 1), more = TRUE)
print(plot_e3b, position = c(0, 0, 1, 0.5))
```

_Solution:_ 

The second distribution is more widely scattered around the mean that, hence, result a little smaller.

### Exercise 4 - 25/3/'16

Have a normal distribution with mean 110 and standard deviation 15. 

```{r 4.1}
e4a <- seq(60, 170, length = 10000)
e4b <- dnorm(e4a, 110, 15)

plot_e4a <- xyplot(e4b ~ e4a,
       type = "l",
       main = "Plot 1",
       scales = list(x = list(at = seq(60, 170, 10)), rot = 45),
       panel = function(x, ...){
               panel.xyplot(x, ...)
               panel.abline(v = 110, lty = 2)
       })

plot_e4b <- xyplot(e4b ~ e4a,
       type = "l",
       main = "Plot 2",
       scales = list(x = list(at = seq(60, 170, 10)), rot = 45),
       panel = function(x,y, ...){
               panel.xyplot(x,y, ...)
               panel.abline( v = c(110, 125), lty = 2)
               
               xx <- c(110, x[x>=110 & x<=125], 125) 
               yy <- c(0,   y[x>=110 & x<=125], 0) 
               panel.polygon(xx,yy, ..., col='red')
       })

plot_e4c <- xyplot(e4b ~ e4a,
       type = "l",
       main = "Plot 3",
       scales = list(x = list(at = seq(60, 170, 10)), rot = 45),
       panel = function(x,y, ...){
               panel.xyplot(x,y, ...)
               panel.abline( v = c(95, 110, 140), lty = 2)
               
               xx <- c(95, x[x>=95 & x<=140], 140) 
               yy <- c(0,   y[x>=95 & x<=140], 0) 
               panel.polygon(xx,yy, ..., col='red')
       })

plot_e4d <- xyplot(e4b ~ e4a,
       type = "l",
       main = "Plot 4",
       scales = list(x = list(at = seq(60, 170, 10)), rot = 45),
       panel = function(x,y, ...){
               panel.xyplot(x,y, ...)
               panel.abline( v = c(80, 95, 110), lty = 2)
               
               xx <- c(85, x[x>=80 & x<=95], 95) 
               yy <- c(0,   y[x>=80 & x<=95], 0) 
               panel.polygon(xx,yy, ..., col='red')
       })

print(plot_e4a, split = c(1, 1, 2, 2), more = TRUE)
print(plot_e4b, split = c(2, 1, 2, 2), more = TRUE)
print(plot_e4c, split = c(1, 2, 2, 2), more = TRUE)
print(plot_e4d, split = c(2, 2, 2, 2), more = FALSE)

```

__A. About what percentage of the values lie between 110 and 125?__

Between 110 and 125 we have approximately the 34% of the observations.

__B. About what percentage of the values lie between 95 and 140?__

Between 95 and 140 there are approximately the 81% of the observations.

__C. About what percentage of the values lie between 80 and 95?__ 

Between 80 and 95 we should expect about the 12% of the observations.

### Exercise 5 - 30/3/'16

Exam scores have a normal distribution with mean of 70 and standard deviation of 10. Bob's score is 80. Find and intepret his standard score. 

```{r 5.1}
e5a <- seq(30, 110, length = 1000)
e5b <- dnorm(e5a, 70, 10)

e5c <- seq(-3, 3, length = 1000)
e5d <- dnorm(e5c, mean = 0, sd = 1)

# Bob's Z-score:
(80 - 70)/10
```

__Solution:__ 

Bob's Z-score is 1, that is, exactly one standard deviation from the mean.

```{r 5.2 plots}

plot_e5a <- xyplot(e5b ~ e5a, 
                  type = "l",
                  main = "Bob's score",
                  panel = function(x, ...){
                          panel.xyplot(x, ...)
                          panel.abline(v = c(70, 80), lty = c(1, 2))
                          panel.text(lab = c("Mean", "Bob"), x = c(70, 80), y = 0.02, 
                adj = c(0.4, 0), cex = 1.5, srt = 45)
                  })


plot_e5b <- xyplot(e5d ~ e5c, 
                  type = "l",
                  main = "Bob's standardized score",
                  panel = function(x, ...){
                          panel.xyplot(x, ...)
                          panel.abline(v = c(0, 1), lty = c(1, 2))
                          panel.text(lab = c("Mean", "Bob"), x = c(0, 1), y = 0.2, 
                adj = c(0.4, 0), cex = 1.5, srt = 45)
                  })

print(plot_e5a, split = c(1, 1, 1, 2), more = TRUE)
print(plot_e5b, split = c(1, 2, 1, 2), more = FALSE)
dev.off()
```

### Exercise 6 - 4/4/'16

Bob scores 80 on both his math exam (which has a mean of 70 and standard deviation on 10) and his English exam (which has a mean of 85 and a standard deviation of 5). Find and interpret Bob's z-scores on both exams to let him know wich exam (if either) he did better on.

```{r 6.1 dataload}
e6a <- seq(0, 100, length = 1000)
e6b <- dnorm(e6a, mean = 70, sd = 10)

e6c <- seq(0, 100, length = 1000)
e6d <- dnorm(e6c, mean = 85, sd = 5)
```

```{r 6.2 plots}
plot_e6a <- xyplot(e6b ~ e6a,
                   type = "l",
                   main = "Math exam",
                   ylim = range(range(e6d)),
                   panel = function(x, ...){
                           panel.xyplot(x, ...)
                           panel.abline(v = c(70, 80), lty = c(2, 3))
                           panel.text(lab = c("Mean", "Bob"), x = c(70, 80), y = 0.03, 
                adj = c(0.2, 0), cex = 1.5, srt = 90)
                           })

plot_e6b <- xyplot(e6d ~ e6c,
                   type = "l",
                   main = "English exam",
                   panel = function(x, ...){
                           panel.xyplot(x, ...)
                           panel.abline(v = c(85, 80), lty = c(2, 3))
                           panel.text(lab = c("Mean", "Bob"), x = c(85, 80), y = 0.03, 
                adj = c(0.4, 0), cex = 1.5, srt = 90)
                           })
print(plot_e6a, split = c(1, 1, 1, 2), more = TRUE)
print(plot_e6b, split = c(1, 2, 1, 2), more = FALSE)
```

Now lets draw the standard normal distibution of the exams and the standardized results of Bob. For the math exam, the standardized score for Bob is `r (80-70)/10` and `r (80-85)/5` for the other. 
Bob had a worst result in the second exam.

```{r 6.3 plot standard normal}
e6e <- seq(-3, 3, length = 1000)
e6f <- dnorm(e6e, 0, 1)

plot_e6c <- xyplot(e6f ~ e6e, 
                   type = "l",
                   main = "Math standardized scores",
                   panel = function(x, ...){
                           panel.xyplot(x, ...)
                           panel.abline(v = 0, lty = 2)
                           panel.text(lab = "Mean and Bob", x = 0, y = 0.15, adj = c(0.2, -0.2), cex = 1.2, srt = 45)
                   })

plot_e6d <- xyplot(e6f ~ e6e, 
                   type = "l",
                   main = "English standardized scores",
                   panel = function(x, ...){
                           panel.xyplot(x, ...)
                           panel.abline(v = c(0, -1), lty = c(2, 3))
                           panel.text(lab = c("Mean", "Bob"), x = c(0, -1 ), y = 0.25, adj = c(0.4, -0.2), cex = 1.3, srt = 90)
                   })

print(plot_e6c, split = c(1, 1, 1, 2), more = TRUE)
print(plot_e6d, split = c(1, 2, 1, 2), more = FALSE)
```

### Exercise 7 - 5/4/'16

Sue's math class has a mean of 70 with a standard deviation of 5. Her standard score is -2. What is her original score?

```{r 7.1 dataload}
e7a <- seq(-3, 3, length = 1000)
e7b <- dnorm(e7a, 0, 1)
e7c <- seq(0, 100, length = 1000)
e7d <- dnorm(e7c, 70, 5)

plot_e7a <- xyplot(e7b ~ e7a,
                   type = "l",
                   main = "Sue's standard math score",
                   panel = function(x, ...){
                           panel.xyplot(x, ...)
                           panel.abline(v = c(0, -2), lty = c(2, 3))
                           panel.text(lab = c("Mean", "Sue"), x = c(0, -2), y = 0.03, adj = c(0.4, 0), cex = 1.2, srt = 45)
                   })

plot_e7b <- xyplot(e7d ~ e7c,
                   type = "l",
                   main = "Sue's math result",
                   scales = list(y = list(at = seq(0, 0.4, 0.1))),
                   panel = function(x, ...){
                           panel.xyplot(x, ...)
                           panel.abline(v = c(70, 60), lty = c(2, 3))
                           panel.text(lab = c("Mean", "Sue"), x = c(70, 60), y = 0.03, adj = c(0.4, 0), cex = 1.2, srt = 45)
                   })

print(plot_e7a, split = c(1, 1, 1, 2), more = TRUE)
print(plot_e7b, split = c(1, 2, 1, 2), more = FALSE)
```

### Exercise 8 - 6/4/'16

Suppose your score on the exam is directly at the mean. What is your standard score.

```{r 8.1 dataload}
e8a <- seq(-3, 3, length = 1000)
e8b <- dnorm(e8a, 0, 1)
```

```{r 8.2 plot standard normal}
xyplot(e8b ~ e8a,
       type = "l",
       main = "Standard normal distribution",
       panel = function(x, ...){
               panel.xyplot(x, ...)
               panel.abline(v = 0, lty = 2)
               panel.text(lab = "Mean", x = 0, y = 0.2, adj = c(0.4, 0.2), cex = 1.2, srt = 45)
       })

```

**Solution:**

The mean of a standard normal distribution is always zero. And since the z-score is $(x-\mu)/\sigma$, it would result in 0.

### Exercise 9 - 6/4/'16

SUppose the weight of cereal boxes have a normal distribution with mean of 20 ounces and a standard deviation of half an ounce. A box that has a standard score of zero weighs how much?

**Solution:**

The standard score is $z = (x-\mu)/\sigma$. In this case, a cereal box with standard score of zero weighs $x = (z/\sigma)-\mu$, that is 20 ounces.

### Exercise 10 - 6/4/'16

Suppose you want to put fat Fido on a weight-loss program. Before the program, his weight had a standard score of +2 compared to the dogs of its breed/age, and after the program, his weight has a standard score of -2. His weight before the program was 150 pounds, and the standard deviation for the breed is 5.

```{r 10.1 data load}
e10a <- seq(-4, 4, length = 1000)
e10b <- dnorm(e10a, 0, 1)

xyplot(e10b ~ e10a, type = "l",
       main = "Fido",
       panel = function(x, ...){
               panel.xyplot(x, ...)
               panel.abline(v = c(2, -2), lty = 3)
               panel.text(lab = c("Before", "After"), x = c(2, -2), y = 0.2, adj = c(0.2, 0), cex = 1.3, srt = 45)
       })
```

**A. What is the mean weight for Fido's breed/age?**

Since $z = (x - \mu)/\sigma$, the breed mean $\mu$ is $-z\sigma + x$, that is `r -2*5 + 150`. 

**B. What is its weight after the weight-loss program?**

After the program the weight $x_2$ is $ \mu - 2\sigma$, that is `r 140 - 2*5`.

### Exercise 11 - 12/4/'16 - First EasieR duty!

Bob's commuting times to work hava normal distribution with a mean of 45 minutes and a standard deviation of 10 minutes. 

**A. What percentage of time does bob get to work in 30 minutes or less?**

```{r 11.1}
z_plot(30, 45, 10, p = 1)
```

Since $z$ is $(X - \mu)/\sigma$ is -1.5, the probability for Bob to get to work in 30 minutes or less is `r pnorm(30, 45, 10)`, that is 6.7%.

**B. Bob's workday starts at 9am, if he leaves at 8am, how often he is late?**

```{r 11.2}
z_plot(60, 45, 10, 2)
```

The `r 1-pnorm(60, 45, 10)`%, approximately the 7%.

### Exercise 12 - 12/4/'16 - Easier on duty!

Times to complete a statistics exam have a normal distribution with mean 40 minuteas and standard deviation of 6 minutes. Deshawn's time comes in the 90th percentile. What percentage of the students are still working on their exams when Deshawn leaves?

**Solution:** 10%. 


### Exercise 13 - 15/4/'16 - Easier on duty!

Suppose your exam score has a standard score of 0.9. Does this mean that 90% of the other exam scoes are lower than yours?

**Solution:** No. It means that my score is 0.9 standard daviation above the mean. Because of the empirical rule, I can say that I am approximately part of 65% of the scores of this exam, since the 68% of scores usually are between one standard deviation from the mean. Futhermore, a z-score of 0.9 corresponds to the 81.59th percentile.

### Exercise 14 - 15/4/'16 - Easier on duty!

If a beby's weight is ad the median, what is her percentile?

**Solution:** It is the 50th percentile.

### Exercise 15 - 15/4/'16 - EasieR on duty!

Clint sleeps an average of 8 hours per night with a standard deviation of 15 minutes. What is the chance he will sleep less than 7.5 hours tonight?

```{r 15.1}
z_plot(7.5, 8, 0.25, 1)

```

**Solution:** 

The probability for Clint to sleep less than 7.5 hours is 2.3%.

### Exercise 16 - 15/4/'16 - EasieR on duty!

Suppose you know that Bob's test score is above the mea, but he does not remember how much. At least, how many students must score lower than Bob?

**Solution:**

At least the 50%.

### Exercise 17 - 15/4/'16 - EasieR on duty!

Bob's commuting times to work have a normal distribution with a mean of 45 minutes and a standard deviation of 10 minutes. How often does Bob get to work in 30 to 45 minutes?

```{r 17.1}
z_plot(30, 45, 10, 3, 45)
```

**Solution:**

Bob gets to work in 30 to 45 minutes the ~43% of the times.

### Exercise 18 - 15/4/'16 - EasieR on duty!

The times taken to complete a statistics exam has a normal distribution with a mean of 40 minutes and a standard deviation of 6 minutes. What is the chance of Deshawn completing the exam in 30 to 35 minutes?

```{r 18.1}
z_plot(30, 40, 6, p = 3, y = 35)
```

**Solution:** 

The probability for Deshawn of completing the exam in 30 to 35 minutes is 15.4%.