---
title: 'Rumsay, Chapter 6: Measures of relative standing and the normal distribution.
  [Ongoing]'
author: "Maurizio Murino"
date: "12 March 2016"
output: html_document
---

```{r setoption, cache=TRUE, warning=FALSE, message=FALSE, fig.width=12}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE, fig.width=12)
```

```{r, loadlattice}
library(lattice)
```

### Exercise 1 - 12/3/'16

What do you guess are the standard deviations of the two distributions in the figure below?

```{r 1.1 dataload1, echo=FALSE}

set.seed(123)
e1a <- seq(10, 70,length = 10000)
e1b <- dnorm(e1a, mean = 40, sd = 8)
```

```{r plot 1}
plote1 <- xyplot(e1b ~ e1a,
       type = "l",
       horizontal = FALSE,
       scales = list(y = list(at = NULL)),
       main = "Normal distribution, mean: 40, ds: 8, range: 10-70",
       panel = function(x, ...) {
               panel.xyplot(x, ...)
               vlines <- c(32, 40, 48)
               panel.abline( v = vlines, lty = 3)
       })
```     

```{r dataload 2, echo=FALSE}
set.seed(123)
e1c <- seq(10, 70,length = 10000)
e1d <- dnorm(e1c, mean = 40, sd = 4)
```

```{r plot2}
plote2 <- xyplot(e1d ~ e1c,
       type = "l",
       horizontal = FALSE,
       scales = list(y = list(at = NULL)),
       main = "Normal distribution, mean: 40, ds: 4, range: 10-70",
       panel = function(x, ...) {
               panel.xyplot(x, ...)
               vlines <- c(36, 40, 44)
               panel.abline( v = vlines, lty = 3)
       })

## position: (a, b, c, d) b: height from the top; d: height from the bottom

print(plote1, position=c(0, .5, 1, 1), more=TRUE)
print(plote2, position=c(0, 0, 1, .5))
```

__Solution:__ 

In the second plot, the standard deviation is slighly smaller than the first's.

### Exercise 2 - 15/3/'16

Draw a picture of a normal distribution with mean 70 and standard deviation 5.

```{r2.1  data load, echo=FALSE}

e2a <- seq(0, 140, length = 10000)
e2b <- dnorm(e2a, 70, 15)                  
```

```{r2.2 plot}
xyplot(e2b ~ e2a,
       type = "l",
       main = "Normal distribution, mean 70",
       scales = list(y = list(at = NULL), x = list(at = seq(0, 140, 10))),
       panel = function(x, ...){
               panel.xyplot(x, ...)
               vmean <- 70
               panel.abline(v = vmean, lty = 4)
               
       })
```

### Exercise 3 - 24/3/'16

Draw a picture with two normal distributions. Give each a mean of 70, with standard deviations of 5 and 10. How do the distributions differ?

```{R 3.1 plots}
e3a <- seq(40, 100, length = 1000)
e3b <- dnorm(e3a, 70, 5)
plot_e3a <- xyplot(e3b ~ e3a,
                   type = "l",
                   main = "Plot 1 - Mean 70, sd 5")

e3c <- seq(40, 100, length = 1000)
e3d <- dnorm(e3a, 70, 10)
plot_e3b <- xyplot(e3d ~ e3c,
                   type = "l",
                   main = "Plot 2 - Mean 70, sd 10",
                   ylim = range(e3b))
                   

print(plot_e3a, position = c(0, .5, 1, 1), more = TRUE)
print(plot_e3b, position = c(0, 0, 1, 0.5))
```

_Solution:_ 

The second distribution is more widely scattered around the mean.