---
title: 'Rumsay, Chapter 5: Undestanding probability basis'
author: "Maurizio Murino"
date: "4 March 2016"
output: html_document
---
```{r setoption, cache=TRUE, warning=FALSE, message=FALSE, fig.width=12}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE, fig.width=12)
```

### Exercise 1 - 4/3/'16

Balls colors: "brown" 13%, "yellow" 14%, "red" 13%, "blue" 24%, "orange" 20%, "green" 16%.

```{r 1.1 dataload}
# There is no sample dimension. I make a n = 100 dataframe.

e3 <- data.frame(color = rep(c("brown", "yellow", "red", "blue", "orange", "green"), c(13, 14, 13, 24, 20, 16)))
barchart(~table(color), data = e3,
         xlab = "Percentage",
         main = "Balls (no N)")
```

__A. What is the chance that you extract a brown or yellow ball?__

It is the sum of the quantities of yellow and brown balls on the total, that is:

```{r 1.2 prob}
paste((13 + 14), "%", sep = "")
```

__B. What is the chance that you will not pull out a blue?__

Blue balls represent the 24% of the pool. Hence, the probability of not extracting it is:

```{r 1.3 prob}
paste((100 - 24), "%", sep = "")
```

### Exercise 2 - 6/3/'16

Flip a coin four times and have four head. Does this outcome give you the idea that the coin is not fair?

__Solution:__

We do not have enough data to answer this question. Four coin tosses produce `2^4 = 16` possible results. Four heads represents a single possible result, hence the probability of getting it is `1/16 = .06`. It is not likely, but it is possible. 

### Exercise 3 - 6/3/'16

Toss a fair coin 10 times, and record how many heads occur.

__A. How many are the possible outcomes?__

```{r 3.1}
2 ^ 10 
```

__B. What would be the probability of each of the outcomes?__

```{r 3.2}
paste(round((1 / 1024), 3), "%", sep = "") 
```

__c. How many of the outcomes would have one head? What is the probability of one head on ten flips.__

A possible solution id HTTTTTTTTTT, but I could have a head at any toss, such as the second THTTTTTTTT and so forth. Hence, we can count 10 possible outcomes:

```{r 3.3}
paste(round((10 / 1024), 2), "%", sep = "") 
```

__D. How many of the outomes would have zero heads? What is the probability of zero heads in ten filps?__

Only one possible result has no heads.

```{r 3.4}
paste(round((1/1024), 3), "%", sep = "")
```

__E. What is the probability of getting one head or less on 10 flips?__

