---
title: 'Rumsay, Chapter 7: Demystifying sampling distributions ans the central limit
  theorem.'
author: "Maurizio Murino"
date: "22 June 2016"
output: html_document
---

```{r setoption, cache=TRUE, warning=FALSE, message=FALSE, fig.width=12}
knitr::opts_chunk$set(cache=TRUE, warning=FALSE, message=FALSE, fig.width=12)
```

```{r, loadlattice}
library(EasieR)
library(knitr)
library(lattice)
library(sn)
```

### Exercise 1 - 22/6/'16

Suppose you take a sample of 100 from a skewed population with mean 50 an standard deviation 15.

**a:** What sample size condition do you need to check here?
**b:** What's the shape and center of the sampling distribution for $\=X$?
**c:** What's the standard error?


**Solution:** 

**a:** The sample should be at least 30. The condition is met.
**b:** Because of the Central Limit Theorem that states that every sample has approximately a normal distribution, we can say that the sample distribution is bell shaped. The center is 50, since the mean of a sample distribution tends to be like the population mean.
**c:** The standard error of the sample distribution is $\sigma/\sqrt{n}$, that is 1.5.
