---
title: 'Rumsey, Chapter 2: Organizing categorial data - charts and graphs. [Ongoing]'
author: "Maurizio Murino"
date: "5 February 2016"
output: html_document
---
### Exercise 1 - 5/2/'16

Suppose 375 individuals are asked what type of vehicle they own: SUV, truck, ora car. See the following frequency table.

```{r}
e1 <- data.frame(category = integer(375)) 
e1$category <- rep(c("SUV","Truck","Car"), c(150,125,100))

# Frequency tables
xtabs(e1) 
```

__A. Make a relative frequency table of these results__

```{r}
round(prop.table(table(e1$category)), digits = 2)
```

__B. Make a pie chart of these results__

```{r chart}
pie(table(e1), labels = paste(c("SUV ","Truck ","Car "), round(prop.table(table(e1$category))*100, digits = 1), "%", sep = "") , main = "Vehicles - n: 375", col = rainbow(3))
legend("topleft", c("SUV","Truck","Car"), cex = 0.6, fill = rainbow(3))
```

C. Interpret the results

The sample of cars (375 vehicles) is composed by trucks, cars and SUVs in the proportions of 40, 33.3 and 26.7 percent respectively.

### Excercise 2 - 5/2/'16

For a month, a restaurant owner checks off whem a customer patronize his restaurant. He records the data for 1000 customers for one month. 

```{r2 chart}
customers <- data.frame(category = integer(1000))
customers$category <- rep(c("Breakfast", "Lunch", "Dinner", "Other"), c(10*6.7, 10*22.2, 10*44.4, 10*26.7))
xtabs(customers)
pie(table(customers), labels = paste(c("Breakfast", "Lunch", "Dinner", "Other"), round(prop.table(table(customers$category))*100, digits = 2), "%" ), main = "Restaurant customers - n: 1000", col = rainbow(4))
legend("topright", c("Breakfast", "Lunch", "Dinner", "Other"), cex = 0.5, fill = rainbow(4))
```

__A. What does this information tell the restaurant owner?__

Its restaurant is particularly appreciated in the midday. The 44.4% of the 1000 sampled customers patronize the restaurant for lunch. That is the breakfast restaurant of choice only for the 6.7% of the customers. Better result for dinner, with 22.2%. A relevant 26.7% of customers is uncategorized.

__B. What is the problem with the "other" category? How can this study improved in the future?__

The category "other" covers more customers than "dinner" and "breakfast". Such an amount of information should be properly investigated. It could group categories of customers that normally do not fit the usual schedules, but that could be enough relevant represent a potential economic gain, if treated adequately. 

### Excercise 3 - 7/2/16

A manager decides to categorize her e-mails into five groups: 1) highest, 2) medium, 3) low priority, 4) personal and 5) SPAM.

```{r ex2.3}
e3 <- data.frame(400)
e3 <- rep( c("high", "medium", "low", "personal", "SPAM"), c(60, 120, 20, 50, 150))

# Frequency table
table(e3)
```

__A. Make a relative frequency table.__

```{r ex2.3a}
paste(prop.table(table(e3))*100, "%", sep = "")
```

__B Make a pie chart of this data.__

```{r ex2.3b}
# Does not work with Knitr.
# par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)

pie(table(e3), labels = paste(round(prop.table(table(e3))*100), "%", sep = ""), col = heat.colors(5), main = "Mail received - n: 400")

# With Knitr the legend is not visible, while there is no problem with the normal plot output.

#usr <- par("usr")*1.2; rect(usr[1], usr[3], usr[2], usr[4], lwd=2, col=NA)
legend("topright", legend = c("High", "Medium", "Low", "Personal", "SPAM"), fill = heat.colors(5), title = "Categories", cex = 0.6)

```

__C. Interpret the result__

Being able of immediately point SPAM can reprensent a relevant contribute to productivity, since they are are the most conspicuous category: 38% of 400.

### Excercise 4 - 10/2/16

Suppose a survey is conducted to see what tyoes od pets people own. The survey of 100 adults finds that 40 of the own a dog, 60 a cat, 20 a fish, 10 a rodent. Can this data be organized in a pie chart?

__Solution:__
It cannot. Since each adult is allowed to possess more than one animal, they can be counted more than one time(40 + 60 + 20 + 10 = 130). Hence, the observed cases would become more than the sample (100). 
In this case, an histogram would be preferable.

```{r 8}
e8 <- data.frame(100)
e8 <- rep(c("dog", "cat", "fish", "rodent"), c(40, 60, 20, 10))
table(e8)
barplot(table(e8), main = "Animals possessed by 100 adults", xlab = "Animal", ylab = "Count", col = gray.colors(4))

```

### Excercise 5 - 10/2/16

A road intersection is observed for four hours, observing if the drivers stop completely, roll through, ran the stop sign. 

``` {r 5.1}
e5 <- data.frame(rep(c("Complete stop", "Rolled through", "Ran the stop sign"), c(64.2, 35.2, 0.6)))
par("mar")
par(mar=c(1,1,1,1))
pie(table(e5), main = "Drivers behaviour", labels = paste(c("Stop ", "Roll through ", "Ran the stop "), round(prop.table(table(e5))*100, 1), "%", sep = ""))
par(oma=c(2,2,2,2), mar=c(4.1, 4.1, 4.1, 4.1), xpd=TRUE) ; box("inner", col="black", lwd=3)
```

__A. Interpret the result as they appear on the chart__

What an unusual place. The 64.6% of the drivers completely stop at the sign. And even if the 35.4% roll throug the sing, only the 0.6% run the sign stop! OMG! That is not Italy. At all.

__B. What information is missing from the pie chart?__

The dimension of the sample is missing.

__C. Does the missing information affect the interpretation of the results?__

Heavily. We cannot derive any information, since this plot could refer to almost any number of drivers.

__D. Should you make a generalization of all drivers based on this data?__

Not at all. Because of the reason stated in point B. we cannot even think any kind of inference.

### Excercise 6 - 27/2/16

A survey is conducted to determine whether 20 office employees of a certain company would prefer to work at home, if given the chance. Of the 10 women surveyed, 7 said yes, and three no. Of the 10 men, 8 said no, 2 said yes. Compare the result using two pie charts. Does gender seems to affect one's preferences to work at home?

```{r 6.1}
rm(e6)
e6 <- data.frame(gender = rep( c("man", "woman"), each=10))
e6$answer <- ifelse(e6$gender == "man", rep( c("yes", "no"), c(8, 2)), rep( c("yes", "no"), c(3, 7)))
```

``` {r 6.2}
par(mfrow = c(1, 2))

suppressWarnings(pie(table(e6$answer[e6$gender == "man"]), main = "Men", line = -4, labels = paste(c("Yes ", "No "), round(prop.table(table(e6$answer[e6$gender == "man"]))*100, 1), "%", sep = "")))

suppressWarnings(pie(table(e6$answer[e6$gender == "woman"]), main = "Women", line = -4, labels = paste(c("Yes ", "No "), round(prop.table(table(e6$answer[e6$gender == "woman"]))*100, 1), "%", sep = "")))

suppressWarnings(mtext("Willingness to work at home", side = 3, line = -2, outer = TRUE, cex=2,font=2))
```


Gender appears to be a relevant variable relatively to the willingness to work at home.