---
title: 'Rumsey, Chapter 2: Organizing categorial data - charts and graphs. [Ongoing]'
author: "Maurizio Murino"
date: "5 February 2016"
output: html_document
---
### Exercise 1 - 5/2/'16

Suppose 375 individuals are asked what type of vehicle they own: SUV, truck, ora car. See the following frequency table.

```{r}
e1 <- data.frame(category = integer(375)) 
e1$category <- rep(c("SUV","Truck","Car"), c(150,125,100))

# Frequency tables
xtabs(e1) 
```

__A. Make a relative frequency table of these results__

```{r}
round(prop.table(table(e1$category)), digits = 2)
```

__B. Make a pie chart of these results__

```{r chart}
pie(table(e1), labels = paste(c("SUV ","Truck ","Car "), round(prop.table(table(e1$category))*100, digits = 1), "%", sep = "") , main = "Vehicles - n: 375", col = rainbow(3))
legend("topleft", c("SUV","Truck","Car"), cex = 0.6, fill = rainbow(3))
```

C. Interpret the results

The sample of cars (375 vehicles) is composed by trucks, cars and SUVs in the proportions of 40, 33.3 and 26.7 percent respectively.

### Excercise 2 - 5/2/'16

For a month, a restaurant owner checks off whem a customer patronize his restaurant. He records the data for 1000 customers for one month. 

```{r2 chart}
customers <- data.frame(category = integer(1000))
customers$category <- rep(c("Breakfast", "Lunch", "Dinner", "Other"), c(10*6.7, 10*22.2, 10*44.4, 10*26.7))
xtabs(customers)
pie(table(customers), labels = paste(c("Breakfast", "Lunch", "Dinner", "Other"), round(prop.table(table(customers$category))*100, digits = 2), "%" ), main = "Restaurant customers - n: 1000", col = rainbow(4))
legend("topright", c("Breakfast", "Lunch", "Dinner", "Other"), cex = 0.5, fill = rainbow(4))
```

__A. What does this information tell the restaurant owner?__

Its restaurant is particularly appreciated in the midday. The 44.4% of the 1000 sampled customers patronize the restaurant for lunch. That is the breakfast restaurant of choice only for the 6.7% of the customers. Better result for dinner, with 22.2%. A relevant 26.7% of customers is uncategorized.

__B. What is the problem with the "other" category? How can this study improved in the future?__

The category "other" covers more customers than "dinner" and "breakfast". Such an amount of information should be properly investigated. It could group categories of customers that normally do not fit the usual schedules, but that could be enough relevant represent a potential economic gain, if treated adequately. 

### Excercise 3 - 7/2/16

A manager decides to categorize her e-mails into five groups: 1) highest, 2) medium, 3) low priority, 4) personal and 5) SPAM.

```{r ex2.3}
e3 <- data.frame(400)
e3 <- rep( c("high", "medium", "low", "personal", "SPAM"), c(60, 120, 20, 50, 150))

# Frequency table
table(e3)
```

__A. Make a relative frequency table.__

```{r ex2.3a}
paste(prop.table(table(e3))*100, "%", sep = "")
```

__B Make a pie chart of this data.__

```{r ex2.3b}
# Does not work with Knitr.
# par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)

pie(table(e3), labels = paste(round(prop.table(table(e3))*100), "%", sep = ""), col = heat.colors(5), main = "Mail received - n: 400")

# With Knitr the legend is not visible, while there is no problem with the normal plot output.

#usr <- par("usr")*1.2; rect(usr[1], usr[3], usr[2], usr[4], lwd=2, col=NA)
legend("topright", legend = c("High", "Medium", "Low", "Personal", "SPAM"), fill = heat.colors(5), title = "Categories", cex = 0.6)

```

__C. Interpret the result__

Being able of immediately point SPAM can reprensent a relevant contribute to productivity, since they are are the most conspicuous category: 38% of 400.

### Excercise 4 - 10/2/16

Suppose a survey is conducted to see what tyoes od pets people own. The survey of 100 adults finds that 40 of the own a dog, 60 a cat, 20 a fish, 10 a rodent. Can this data be organized in a pie chart?

__Solution:__
It cannot. Since each adult is allowed to possess more than one animal, they can be counted more than one time(40 + 60 + 20 + 10 = 130). Hence, the observed cases would become more than the sample (100). 
In this case, an histogram would be preferable.

```{r 8}
e8 <- data.frame(100)
e8 <- rep(c("dog", "cat", "fish", "rodent"), c(40, 60, 20, 10))
table(e8)
barplot(table(e8), main = "Animals possessed by 100 adults", xlab = "Animal", ylab = "Count", col = gray.colors(4))

```

